Values Encoded in Machine Learning Research

### 2025-02-17 latex/paper.tex
final updates

\subsubsection{ICML and Neurips 2023}
In Figure \ref{fig:percentage_both_conferences} we show the prevelance of values in all 200 papers combined from ICML and Neurips 2023. The top values are performance (97.5\%), accuracy (75\%), 
State-of-the-art (71\%), Efficiency (70\%) and Improvement (67.5\%). Among the values related to user rights and stated in 
ethical principles the most common ones are Privacy (16.5\%) and `Explicability' (15.5\%).

@@

By analysing the 100 most influencial paper from both conferences seperately we can see in Figure \ref{fig:percentage_comparison_conferences}
that the distribution of values are quite similar in both conferences. The values `Performance', `Accuracy' and `State-of-the-art' are among the top 5 values for both conferences. 
One can consider that the maximum length for a paper is nine pages for Neurips\footnote{\url{https://neurips.cc/Conferences/2023/PaperInformation/NeurIPS-FAQ}} and only eight pages for ICML\footnote{\url{https://icml.cc/Conferences/2023/StyleAuthorInstructions}}.
This different limitation in the maximum length of the paper does not seem to have a big impact on the distribution of values in the paper.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../plots/percentage_comparison_conferences.png}
    \caption{Comparison of prevalence of values in the 200 most influential papers from ICML 2023 and NeurIPS 2023. Values related to user rights are highlighted in orange, and values belonging to ethical principles are highlighted in blue.}
    \label{fig:percentage_comparison_conferences}
\end{figure}

@@

This paper analyses papers from the two big ML conferences ICML and Neurips. There are specialized ML conferences dedicated to the study of fairness, accountability and transparency in AI. It can be debated whether the fact that these specialized conferences exist ensures that a focus on performance values at ICML and Neurips seems justifiable.

>>>>

\subsubsection{ICML and NeurIPS 2023}
In Figure \ref{fig:percentage_both_conferences} we show the prevelance of values in all 200 papers combined from ICML and NeurIPS 2023. The top values are performance (97.5\%), accuracy (75\%), 
State-of-the-art (71\%), Efficiency (70\%) and Improvement (67.5\%). Among the values related to user rights and stated in 
ethical principles the most common ones are Privacy (16.5\%) and `Explicability' (15.5\%).

@@

By analysing the 100 most influencial paper from both conferences seperately we can see in Figure \ref{fig:percentage_comparison_conferences}
that the distribution of values are quite similar in both conferences. The values `Performance', `Accuracy' and `State-of-the-art' are among the top 5 values for both conferences. 
One can consider that the maximum length for a paper is nine pages for NeurIPS\footnote{\url{https://neurips.cc/Conferences/2023/PaperInformation/NeurIPS-FAQ}} and only eight pages for ICML\footnote{\url{https://icml.cc/Conferences/2023/StyleAuthorInstructions}}.
This different limitation in the maximum length of the paper does not seem to have a big impact on the distribution of values in the paper.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../plots/percentage_comparison_conferences.png}
    \caption{Comparison of prevalence of values in the 200 most influential papers from ICML 2023 and NeurIPS 2023. Values related to user rights are highlighted in orange, and values belonging to ethical principles are highlighted in blue.}
    \label{fig:percentage_comparison_conferences}
\end{figure}

@@

This paper analyses papers from the two big ML conferences ICML and NeurIPS. There are specialized ML conferences dedicated to the study of fairness, accountability and transparency in AI. It can be debated whether the fact that these specialized conferences exist ensures that a focus on performance values at ICML and NeurIPS seems justifiable.
